---
title: "New York City Job Openings!"
author: "Nicholas Stover"
date: "October 25, 2018"
output: 
  html_document:
    theme: readable
    code_folding: hide
---


# Background
New York City is my wife's dream life location! I support this location. I found some data on kaggle.com that has job openings! This will be an exciting project. I want to analyze this data and see what I can find. We are both interested in these results, even if they don't provide much use.



```{r, include=FALSE}
library(readr)
library(tidyverse)
library(ggplot2)
library(stringr)
```

# Clean
The data is very messy. It includes huge texts of job description and other not useful columns. I just want to visualize this data quickly and easily. I will remove some of these columns that don't provide useful information for this project. These texts could definitely be useful later if we find some job openings that might be a fit for us!

Tidyverse is very useful for this.

```{r, echo=FALSE}
nyc_jobs <- read_csv("Data/nyc-jobs.csv") %>% 
  ungroup() %>% 
  select(-`Job ID`, -`Title Code No`, -`Job Description`, -`Minimum Qual Requirements`, -`Preferred Skills`, -`Additional Information`, -`To Apply`, -`Residency Requirement`, -`Recruitment Contact`, -`Post Until`, -`Posting Updated`, -`Process Date`, -`Work Location 1`, -`Work Location`, -`Posting Type`, -Level, -`Salary Frequency`) %>%
  as_tibble()
```

I wanted to create a column with the average salary because the salary range columns aren't too useful.

```{r}
nyc_jobs$avgsalary <- (nyc_jobs$`Salary Range From` + nyc_jobs$`Salary Range To`)/2
```


# Creating Subsets

I created a subset searching for certain keywords in the job titles for both me and my wife.


```{r}
nic_nyc_jobs <- nyc_jobs %>%
  ungroup() %>%
  select(-`Salary Range From`, -`Salary Range To`) %>%
  filter(avgsalary > 10000, nyc_jobs$`# Of Positions` < 100, str_detect(nyc_jobs$`Business Title` , '\\b(Analyst|Scientist|Data|Market|Research|Machine|Learning|Processing|Mining|Statistician|Statistics|Analytics)\\b'))


sadie_nyc_jobs <- nyc_jobs %>%
  ungroup() %>%
  select(-`Salary Range From`, -`Salary Range To`) %>%
  filter(avgsalary > 10000, nyc_jobs$`# Of Positions` < 100, str_detect(nyc_jobs$`Business Title` , '\\b(Design|Stylist|Designer|Textile|Textiles|Fashion|Creative|Buyer)\\b'))
```
# Visualize
The data is a little cleaner to visualize. It only includes the useful columns that might give use some info. I converted agency to factor and converted the "Posting Date" column into dates.

```{r}
nyc_jobs$Agency <- as.factor(nyc_jobs$Agency)
nic_nyc_jobs$`Posting Date` <-    as.Date(nic_nyc_jobs$`Posting Date`)

nic_nyc_jobs <- nic_nyc_jobs %>%
  filter(nic_nyc_jobs$`Posting Date` > as.Date("2018-09-01"))
sadie_nyc_jobs <- sadie_nyc_jobs %>%
  filter(sadie_nyc_jobs$`Posting Date` > as.Date("2018-09-01"))

```

Some visualizations for my job prospects. This is bar chart of the number of job postings per agency.

```{r}
theme_set(theme_classic())
nic_agency <- as.data.frame(table(nic_nyc_jobs$Agency))
colnames(nic_agency) <- c("Agency", "Frequency")
nic_plot <- ggplot(nic_agency, aes(Agency, Frequency))
nic_plot + geom_bar(stat="identity", width = 0.5, fill="blue") + 
  labs(title="Jobs per Agency", 
       caption="Nyc Jobs") +
  theme(axis.text.x = element_text(angle=65, vjust=0.5))
```

The most job postings come from the Department of Environment Protection, Human Rights Commission, and the Office of Payroll Administration.

Due to the small number of postings in my wife's subset of data, visualization isn't needed. However, I decided to use a pie chart because she liked how it looked.


```{r}
sadie_pie <- ggplot(sadie_nyc_jobs, aes(x = "", fill = factor(Agency))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="Agency", 
       x=NULL, 
       y=NULL, 
       title="Pie Chart of Agency", 
       caption="Jobs per Agency") + 
  scale_fill_brewer(palette = "RdPu")

sadie_pie + coord_polar(theta = "y", start=0)

ggplot(sadie_nyc_jobs, aes(x=Agency, y = avgsalary)) +
  theme(axis.text.x = element_text(angle=40, vjust=0.5)) + 
  geom_point(stat="identity", fill = "red")
```


This next graph looks at each agency on the x axis, the average salary on the y axis, and color in terms of the number of positions available. The majority of these jobs have only one opening but a few have more. 

```{r}
ggplot(nic_nyc_jobs, aes(x= Agency, y = avgsalary, color = `# Of Positions`)) +
  theme(axis.text.x = element_text(angle=90, vjust=0.5)) + 
  ylab("Average Salary") + ggtitle("Average Salary by Agency") + 
  geom_point(stat="identity") +
  scale_colour_gradient(low = "orange", high="black")


```

Overall, this information is good to keep around us as we apply for jobs and finish up school. I can target a few positions such as a quality assurance analyst with the fire department because there are a few openings and a decent salary.

My wife has a decent shot within all the agencies because they have relatively the same number of openings. She should shoot for the Department of Transportation or the Department of Parks & Recreation though if she wants to have a higher salary.